### 2011-06-15 docs/SDC/sdc.tex
expand notes to show what else needs to be computed in the network

In the Strang-split version of Maestro, the reaction network solves just
the reaction portion of the species evolution equations along with a
temperature evolution equation.  This is done using a standard ODE solver
(like {\tt VODE}) on the system:
\begin{eqnarray}
\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T) \\
\frac{dT}{dt}   &=& 
    \frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right ).
\end{eqnarray}
Furthermore, we simplify the problem ``freezing'' the
thermodynamics---i.e., $c_p$ is evaluating at the start of the
integration and held constant thereafter.

In the SDC version, the VODE integration at the end of an SDC
iteration is responsible for updating all the thermodynamic quantities
\{$\rho$, $(\rho h)$, and $(\rho X_k$)\}, including both the advective
fluxes (incorporated via a piecewise constant advective flux source
term) and the reactions.  This provides a much stronger coupling between
the physical processes.  In particular, our system now looks like:
\begin{eqnarray}
\frac{d(\rho X_k)}{dt} &=& \rho \omegadot_k(\rho,X_k,T) + A_{\rho X_k} \\
\frac{d(\rho h)}{dt}   &=& \rho \Hnuc + A_{\rho h}
\end{eqnarray}
Here, $A_{\rho X_k}$ and $A_{\rho h}$ are piecewise-constant (in time)
approximations to the change in ${\rho X_k}$ and ${\rho h}$ (respectively)
due to the advection.  These are constructed by calling {\tt density\_advance}
and {\tt enthalpy\_advance} in Maestro and passed into the network solver
during the reaction step.  These terms approximate the following:
\begin{eqnarray}
A_{\rho X_k} &=&  \left [- \nabla \cdot (\rho X_k \Ub) \right ]^{n+1/2} \\
A_{\rho h}   &=&  \left [- \nabla \cdot (\rho h U) + \frac{Dp_0}{Dt} \right ]^{n+1/2}
\end{eqnarray}

@@

The major changes from the non-SDC-enabled burners is the addition of
the advective terms to the system of ODEs, the fact that we integrate
$(\rho X_k)$ instead of just $X_k$, and the need to derive the
temperature from the input state for each RHS evaluation by {\tt
  VODE}.

>>>>

\subsection{Strang-split version}

In the Strang-split version of MAESTRO, the reaction and diffusion
processes operate independent of one-another.  The species and
enthalpy equations are updated through $\Delta t$ using the following
sequence:

\begin{itemize}

\item React for $\Delta t/2$

In the Strang-split version of Maestro, the reaction network solves just
the reaction portion of the species evolution equations along with a
temperature evolution equation.  This is done using a standard ODE solver
(like {\tt VODE}) on the system:
\begin{eqnarray}
\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T) \\
\frac{dT}{dt}   &=& 
    \frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right ).
\end{eqnarray}
Here, $T$ is evolved solely to evaluate the reaction rates,
$\omegadot_k(\rho,X_k,T)$.  Furthermore, we simplify the problem
``freezing'' the thermodynamics---i.e., $c_p$ is evaluating at the
start of the integration and held constant thereafter.

We update the enthalpy at the end of the integration as:
\begin{equation}
h^\mathrm{new} = h^\mathrm{old} + \frac{\Delta t}{2} \Hnuc
\end{equation}

\item Advect for $\Delta t$

Here we solve
\begin{eqnarray}
\frac{\partial(\rho X_k)}{\partial t} &=& 
    -\nabla\cdot(\rho X_k\Ub) \\
\frac{\partial(\rho h)}{\partial t} &=& 
    -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} 
\end{eqnarray}

Note that no explicit reaction terms appear here.  Since the advection
takes place using the state updated from the reaction step, the effect
of the reactions is implicitly contained in the advective update.

\item React for $\Delta t/2$

Finally, we react again, starting with the state left by the advection
step.

\end{itemize}

\subsection{SDC version}
In the SDC version, the VODE integration at the end of an SDC
iteration is responsible for updating all the thermodynamic quantities
\{$\rho$, $(\rho h)$, and $(\rho X_k$)\}, including both the advective
fluxes (incorporated via a piecewise constant advective flux source
term) and the reactions.  This provides a much stronger coupling between
the physical processes.  In particular, our system now looks like:
\begin{eqnarray}
\frac{d(\rho X_k)}{dt} &=& \rho \omegadot_k(\rho,X_k,T) + A_{\rho X_k} \\
\frac{d(\rho h)}{dt}   &=& \rho \Hnuc + A_{\rho h}
\end{eqnarray}
Here, $A_{\rho X_k}$ and $A_{\rho h}$ are piecewise-constant (in time)
approximations to the change in ${\rho X_k}$ and ${\rho h}$ (respectively)
due to the advection.  These are constructed by calling {\tt density\_advance}
and {\tt enthalpy\_advance} in Maestro and passed into the network solver
during the reaction step.  These terms approximate the following:
\begin{eqnarray}
A_{\rho X_k} &=&  \left [- \nabla \cdot (\rho X_k \Ub) \right ]^{n+1/2} \\
A_{\rho h}   &=&  \left [- \nabla \cdot (\rho h U) + \frac{Dp_0}{Dt} \right ]^{n+1/2}
\end{eqnarray}
The construction of the interface states used in the advective terms
uses an iteratively-lagged approximation to the reaction source terms
in the interface prediction.  This explicitly couples the reaction
process to the advection process.

@@

At the end of the integration, we define 
\begin{eqnarray}
I_{\rho X_k} &=& \frac{(\rho X_k)^\mathrm{new} - (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k} \\
I_{\rho h}  &=& \frac{(\rho h)^\mathrm{new} - (\rho h)^\mathrm{old}}{\Delta t} - A_{\rho h}
\end{eqnarray}
These terms are used as the source terms for the advection step in the
next SDC iteration.

\subsection{Summary of Changes}
The major changes from the non-SDC-enabled burners is the addition of
the advective terms to the system of ODEs, the fact that we integrate
$(\rho X_k)$ instead of just $X_k$, and the need to derive the
temperature from the input state for each RHS evaluation by {\tt
  VODE}.
